---
title: "p8105_hw6_yq2251"
author: "TritonD"
date: "11/23/2019"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)
```

# Question 1 Baby birthweight model
```{r}
birthw=read_csv('./data/birthweight.csv')
```

```{r}
bw_tidy=
birthw%>%
  mutate(
  babysex=as.factor(babysex),
  babysex=recode(babysex,"1"="male", "2"="female"),
  frace=as.factor(frace),
  frace=recode(frace,"1"="white", "2"="black","3"="asian", "4"="puerto rican", "8"="other","9"="unknown"),
  malform=as.factor(malform),
  malform=recode(malform,"1"="present", "0"="absent"),
  mrace=as.factor(mrace),
  mrace=recode(mrace, "1"="white", "2"="black","3"="asian", "4"="puerto rican", "8"="other"))
```


```{r}
#check if there is missing value
anyNA(bw_tidy)
```
All results are false, there is no missing value

```{r}
#The general model with all predictors included will be
general_x = lm(bwt ~ ., data = bw_tidy)
summary(general_x)
```
There are 3 variables excluded from the model due to singularities: pnumlbw, pnumsga, wtgain

```{r}
#general model with 3 variables excluded:
Model1 = update(general_x, . ~ . -pnumlbw -pnumsga -wtgain)
summary(Model1)
```
The variable "frace" has "fraceother" with highest p-value 0.953745 (least significant) and other fraces' p value all not significant, we will eliminate it from model.

```{r}
#Model 1 with 1 extra variable excluded:
Model2 = update(Model1, . ~ . -frace)
summary(Model2)
```
The variable "malformpresent" has highest p-value 0.888937 (least significant) and we will eliminate it from model.

```{r}
#Model 2 with 1 extra variable excluded:
Model3 = update(Model2, . ~ . -malform)
summary(Model3)
```
The variable "malform" has highest p-value 0.888937 (least significant) and we will eliminate it from model.

```{r}
#Model 3 with 1 extra variable excluded:
Model4 = update(Model3, . ~ . -malform)
summary(Model4)
```
The variable "ppbmi" has highest p-value 0.759922 (least significant) and we will eliminate it from model.

```{r}
#Model 4 with 1 extra variable excluded:
Model5 = update(Model4, . ~ . -ppbmi)
summary(Model5)
```
The variable "momage" has highest p-value 0.530319 (least significant) and we will eliminate it from model.

```{r}
#Model 5 with 1 extra variable excluded:
Model6 = update(Model5, . ~ . -momage)
summary(Model6)
```
The variable "menarche" has highest p-value 0.245327 (least significant) and we will eliminate it from model.

```{r}
#Model 6 with 1 extra variable excluded:
Model7 = update(Model6, . ~ . -menarche)
summary(Model7)
```
The variable "fincome" has highest p-value 0.068844 (least significant) and we will eliminate it from model.

```{r}
#Model 7 with 1 extra variable excluded:
Model8 = update(Model7, . ~ . -fincome)
summary(Model8)
```
Although "mracesuan" has insignificant p-value, but the other two mrace categories are significant, thus I will keep this variable in the final model.


```{r}
final_model = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mheight + parity + ppwt + smoken + mrace, data = bw_tidy)

#show my final model's variable estimates in a table
final_model %>% 
  broom::tidy() %>% 
  knitr::kable()
final_model %>% 
  broom::glance()
```
The R-squared value is 0.718 for my model, which indicates 71.8% of the variance can be explained by this final model.

# Visualize my model 
```{r}
bw_tidy %>% 
modelr::add_residuals(final_model) %>% 
modelr::add_predictions(final_model) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()+
  geom_hline(yintercept = 0, color = "red") +
  labs(
    title = 'Model predictted values against residuals',
    x = 'Predicted values',
    y = 'Residuals'
  )
```
Most of the predicted values are ranging 2000-4000, slightly left skewed with few outlines. The residuals ranging 1000- (-1000).Red line was shown at y intercept=0

Build example Model 
```{r}
Exp_M1 = lm(bwt ~ blength + gaweeks, data = bw_tidy)
Exp_M1 %>% 
  broom::tidy()
```


Build example Model 2
```{r}
Exp_M2 = lm(bwt ~ babysex + bhead + blength + babysex*bhead + babysex*blength + bhead*blength+babysex*bhead*blength, data = bw_tidy)
Exp_M2 %>% 
  broom::tidy()
```

#Cross validation
```{r}
cv_df = 
  crossv_mc(bw_tidy, 100) %>%
   mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  
  mutate(final_model = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mheight + parity + ppwt + smoken + mrace, data = .x)),
         Exp_M1= map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         Exp_M2= map(train, ~lm(bwt ~ babysex + bhead + blength + babysex*bhead + babysex*blength + bhead*blength+babysex*bhead*blength, data=.x))) %>% 
  
  mutate(rmse_final_model = map2_dbl(final_model, test, ~rmse(model = .x, data = .y)),
         rmse_Exp_M1= map2_dbl(Exp_M1, test, ~rmse(model = .x, data = .y)),
         rmse_Exp_M2 = map2_dbl(Exp_M2, test, ~rmse(model = .x, data = .y)))
```

shows the distribution of RMSE values for each candidate model
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
According to violin plot, my final model fits the data best with least error.



# Question 2
Load the data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
bootstraps_weather=
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy),
    glance = map(models, broom::glance)) %>% 
  select(-models, -strap) %>% 
  unnest(results,glance)%>%
  janitor::clean_names()
```


```{r}
weather_r2 = 
  bootstraps_weather %>% 
  select(id, r_squared)
quantile(weather_r2$r_squared, 0.025)
quantile(weather_r2$r_squared, 0.975)
```
The 95% confidence interval for R-squared is (0.894, 0.927).


```{r}
weather_betas = 
 bootstraps_weather %>% 
  group_by(id) %>% 
  summarize(
    boot_betas = log(estimate[[1]]*estimate[[2]])
  )
quantile(weather_betas$boot_betas, 0.025)
quantile(weather_betas$boot_betas, 0.975)
```
The 95% confidence interval for log(beta0*beta1) is (1.965, 2.061).


# Plot estimated distributions
```{r}
weather_r2 %>% 
  ggplot(aes(x = r_squared)) +
  geom_density()+
  ggtitle("Distribution of R-squared")
```
R-squared has a slightly left-skewed distribution


```{r}
weather_betas %>% 
  ggplot(aes(x = boot_betas)) +
  geom_density()+
  ggtitle("Distribution of log(beta0*beta1)")
```
log(beta0*beta1) has a aproximately normal distribution
